---
title: "GV300 Showcase"
description: "Example code + visualisation work from GV300"
output: distill::distill_article
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)

# Load your dataset
kidiq <- read.csv("kidiq.csv")
```

## Introduction

This page presents a data visualisation from a previous GV300 assignment. 
The aim of the analysis was to examine how maternal characteristics influence children's test scores.

Below is an example of the code used and the resulting visualisation.


```{r supermoms-plot, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

kidiq_plot <- kidiq %>%
  filter(!is.na(work_status), !is.na(kid_score), !is.na(family_advantage), !is.na(iq_group)) %>%
  mutate(
    work_status = factor(work_status),
    family_advantage = factor(family_advantage),
    iq_group = factor(iq_group)
  )

overall_mean <- mean(kidiq_plot$kid_score, na.rm = TRUE)

your_plot <- ggplot(kidiq_plot, aes(x = work_status, y = kid_score, fill = family_advantage)) +
  geom_violin(trim = FALSE, alpha = 0.85, color = "grey30") +
  geom_boxplot(width = 0.12, outlier.shape = NA, alpha = 0.35, color = "grey20") +
  geom_hline(yintercept = overall_mean, linetype = "dashed", linewidth = 0.5) +
  facet_wrap(~ iq_group) +
  scale_fill_brewer(palette = "RdYlBu", direction = -1) +
  labs(
    title    = "SuperMoms: Maternal Advantage",
    subtitle = "Kids thrive when mothers combine education and steady work,\nrevealing how opportunity compounds across generations.",
    x        = "Mother's work status",
    y        = "Child's test score",
    fill     = "Family advantage",
    caption  = "Note: dashed line shows the overall mean child score."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position  = "bottom",
    plot.title       = element_text(face = "bold", size = 16),
    plot.subtitle    = element_text(size = 12),
    axis.text.x      = element_text(angle = 20, hjust = 1),
    strip.text       = element_text(face = "bold")
  )

your_plot
```

- The code first calculates the overall mean child test score and stores it as a reference value. 
- It then creates a violin plot showing the distribution of children’s test scores across different categories of maternal work status, with colours representing levels of family advantage. 
- The violin shapes display the full distribution of scores within each group, allowing us to see variation, density, and spread rather than just averages. 
- A dashed horizontal line marks the overall mean score, making it easier to compare each group to the sample average. 
- The plot is further divided into panels by maternal IQ group, which allows us to examine whether patterns differ across levels of maternal cognitive ability.
- The figure shows that children from more advantaged families generally tend to achieve higher test scores, and that work status appears to interact with this advantage. 
- Differences across IQ groups suggest that while maternal ability may matter, structural factors such as education, employment, and family resources also play a significant role. 
- Overall, the visualisation highlights how socioeconomic advantages can accumulate across generations.