---
title: "Text Analysis: Understanding Political Language"
description: "A simple text analysis exploring word frequency and sentiment in political-style text."
output: distill::distill_article
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidytext)
library(textdata)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Overview

In this section, I demonstrate a basic text analysis using R. The goal is to explore how language is used in political and economic statements by identifying the most frequent words and analysing sentiment.
Text analysis is important in political economy because it allows researchers to systematically study narratives, framing, and discourse in speeches, media, and policy documents.
Data
For this example, I use a small sample of political-style statements:

```{r}
text_data <- tibble(text = c(
  "Economic growth improves living standards and reduces poverty",
  "Inequality remains a major challenge in modern economies",
  "Government policy should promote stability and development",
  "Financial crises create instability and unemployment",
  "Strong institutions support democracy and economic growth"
))
```

## Word Frequency Analysis

First, I clean and break the text into individual words, removing common stopwords such as 'the' and 'and'.

```{r}
tidy_words <- text_data %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)

word_counts <- tidy_words %>%
  count(word, sort = TRUE)

word_counts
```

## Visualisation

```{r}
ggplot(word_counts %>% top_n(10), 
       aes(x = reorder(word, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Most Common Words in Political Text",
    subtitle = "Word frequency after removing stopwords",
    x = "Words",
    y = "Frequency"
  ) +
  theme_minimal()
```

## Sentiment Analysis

Next, I analyse whether the language used is generally positive or negative.

```{r}
sentiment <- tidy_words %>%
  inner_join(get_sentiments("bing")) %>%
  count(sentiment)

sentiment
```


## Visualisation

```{r}
ggplot(sentiment, aes(x = sentiment, y = n)) +
  geom_col() +
  labs(
    title = "Sentiment Analysis of Text",
    subtitle = "Positive vs Negative Word Counts",
    x = "Sentiment",
    y = "Count"
  ) +
  theme_minimal()
```


## Interpretation

The results show that words such as “growth”, “economic”, and “stability” appear frequently, suggesting a strong focus on economic performance and institutional strength.
The sentiment analysis indicates that the language is generally balanced but leans slightly positive, reflecting how political and economic discourse often emphasises progress, development, and stability.
This demonstrates how text analysis can be used to uncover underlying themes and tone in political communication.

